Safe and Robust Formal Synthesis of Digital Controllers\\ for Continuous Plants

### 2018-05-08 automatica/paper.tex
sections 2-3

%% \addtodo{CD: the novelty compared to CAV'17 needs to be clarified}
This paper is an extension of~\cite{DBLP:conf/cav/AbateBCCDKKP17}
and makes the following contributions, where the fourth and the fifth are
extensions over~\cite{DBLP:conf/cav/AbateBCCDKKP17}.
%
\begin{enumerate}

  % Additions from CAV:
  % continuous time
  % floating point 
\item We automatically generate \emph{correct-by-construction} digital
  controllers using an inductive synthesis approach.  Our application of program
  synthesis is non-trivial and addresses challenges specific to control
  systems, such as quantizers and FWL.  %% In particular, we
  %% have found that a two-stage verification engine that continuously refines
  %% the precision of the fixed-point representation of the plant yields a
  %% speed-up of two orders of magnitude over a conventional one-stage
  %% verification engine.
%
  %\item  We compute
  The computed state-feedback controllers 
  guarantee a given safety property alongside stability and safety
  specifications. Existing methods for controller synthesis rely on transfer
  function representations, which are inadequate to prove safety requirements.
%
\item We provide two novel algorithms: the first, na\"ive one, relies on
  an unfolding of the dynamics up to a completeness threshold, while the
  second one is abstraction-based and leverages abstraction refinement and
  acceleration to improve scalability while retaining soundness.  %% A third
  %% algorithm evaluates systems using transfer functions where no state-space
  %% description is given.
  Both approaches provide sound synthesis of
  feedback controllers and consider the various sources of imprecision in
  the implementation of the control algorithm and in the modeling of the
  plant.
%
\item We develop a model for different sources of quantization error and
  their effect on reachability properties.  We give bounds that ensure the
  safety of our controllers in a hybrid continuous-digital domain.

\item While~\cite{DBLP:conf/cav/AbateBCCDKKP17} is restricted to systems
  modeled in discrete time, in the current paper we present a model for
  soundly evaluating reachability in a combined continuous-discrete time/space.

\item A further limitation of the work in~\cite{DBLP:conf/cav/AbateBCCDKKP17}
  is its restriction to fixed-point arithmetic, meaning that fixed-point
  numbers were employed for the representation of both the plant and the
  controller, as well as for the operations performed by each of them. 
  Conversely, in the current paper we also make use of floating-point
  arithmetic and compare the two approaches.  Additionally, we enhance the
  implementation with an SMT backend, allowing us to also use reals in order
  to represent the states of the plant.

@@

\textbf{CEGIS --} Program synthesis is the problem of computing
correct-by-design programs from high-level specifications. Algorithms
for this problem have made substantial progress in recent years, in
particular CEGIS, which is a recent approach to inductive synthesis
borrowing ideas from Counter-Example Guided Abstraction Refinement
(CEGAR)~\cite{DBLP:conf/cav/ClarkeGJLV00} and
superoptimisation~\cite{DBLP:conf/iclp/BrainCVF06,
  DBLP:conf/asplos/Massalin87}.  This approach uses an iterative
process where each iteration performs inductive generalisation based
on counterexamples provided by a verification oracle.  In the absence
of complete specifications, oracles are queried to distinguish between
non-equivalent
programs~\cite{jha-icse10,DBLP:conf/fmcad/AlurBJMRSSSTU13}.

@@

Program synthesizers are an ideal fit for the synthesis of digital controllers, since
the semantics of programs capture the effects of finite-precision arithmetic
precisely.  In~\cite{DBLP:conf/cdc/RavanbakhshS15}, the authors use CEGIS
for the synthesis of switching controllers for stabilizing continuous-time
plants with polynomial dynamics.  The work extends to affine systems, but is
limited by the capacity of the state-of-the-art SMT solvers for solving
linear arithmetic.  Since this approach uses switching models instead of
linear dynamics for the digital controller, it avoids problems related to
finite precision arithmetic, but potentially suffers from state-space
explosion.  Moreover, in \cite{DBLP:conf/emsoft/RavanbakhshS16} the same
authors use a CEGIS-based approach for synthesizing continuous-time
switching controllers that guarantee \emph{reach-while-stay} properties of
closed-loop systems, i.e., properties that specify a set of goal states and
safe states (constrained reachability).  This solution is based on
synthesizing control Lyapunov functions for switched systems that yield
switching controllers with a guaranteed minimum dwell time in each mode. 
However, both approaches are unsuitable for the kind of control we seek to
synthesize.

The work in~\cite{hscc-paper} synthesizes stabilizing controllers for
continuous plants given as transfer functions by exploiting bit-accurate
verification of software-implemented digital controllers~\cite{Bessa16}. 
While this work also uses CEGIS, the approach is restricted to digital
controllers for stable closed-loop systems given as transfer function
models: this results in a static check on their coefficients.  By contrast,
in the current paper we consider a state-space representation of the
physical system, which requires ensuring the specification over actual
traces of the model, alongside the numerical soundness required by the
effects of discretisation and finite-precision errors.  A~state-space model
has known advantages over the transfer function
representation~\cite{Franklin15}: it naturally generalizes to multivariate
systems (i.e., with multiple inputs and outputs); and it allows synthesis of
control systems with guarantees on the internal dynamics, e.g., to
synthesize controllers that make the closed-loop system \emph{safe}.  Our
work focuses on the \emph{safety} of internal states, which is usually
overlooked in the literature.  Moreover, our work integrates an
abstraction/refinement (CEGAR) step inside the main CEGIS loop.

The tool Pessoa~\cite{mazo2010pessoa} synthesizes correct-by-design embedded
control software in a Matlab toolbox.  It is based on the abstraction of a
physical system to an equivalent finite-state machine and on the computation
of reachability properties thereon.  Based on this safety specification,
\mbox{Pessoa} can synthesize embedded controller software for a range of
properties.  The embedded controller software can be more complicated than
the state-feedback control we synthesize, and the properties available cover
more detail.  However, relying on state-space discretization \mbox{Pessoa}
is likely to incur in scalability limitations.  Along this research line,
\cite{Anta2010,liu16} studies the synthesis of digital controllers for
continuous dynamics, and \cite{zamani2014} extends the approach to the
recent setup of Network Control Systems.

\textbf{Discretization Effects --}
Classical work in control synthesis has often disregarded digitalization effects, 
\note{[this paragraph needs more work:]
whereas recent results in digital control have focused on separate aspects of discretization, 
including delayed response~\cite{Duggirala2015} and finite word length (FWL) semantics, 
with the goal either to verify~\cite{daes20161} or to optimize~\cite{oudjida2014design} given implementations. }

@@

For the majority of our benchmarks, we observe that the abstraction-based
back-end is faster than the basic multi-staged verification approach, and
finds one solution more (9) than the multi-staged back-end (8).  In direct
comparison, the abstraction-based approach is on average able to find a
solution in approximately 70\% of the time required using the multi-staged
back-end, and has a median run-time 1.4\,s, which is seven times smaller
than the multi-staged approach.  The two back-ends complement each other in
benchmark coverage and together solve all benchmarks in the set.  On average
our engine spent 52\% in the synthesis and 48\% in the verification phase.

\begin{table*}
\centering
\begin{tabular}{| r | l | c | c | r | c | r | c | r | c | r |}
%
\hline
\# & \multicolumn{1}{|c|}{Benchmark} & \multicolumn{1}{|c|}{Order} & \multicolumn{4}{|c|}{Multi-staged}                 & \multicolumn{4}{|c|}{Abstraction} \\
   &                                  & & \multicolumn{1}{|c|}{$\mathcal{F}_{\langle I_p,F_p \rangle}$} & \multicolumn{1}{|c|}{Time} & \multicolumn{1}{|c|}{$\mathcal{F}_{\langle p \rangle}$} & \multicolumn{1}{|c|}{Time} & \multicolumn{1}{|c|}{$\mathcal{F}_{\langle I_p,F_p \rangle}$} & \multicolumn{1}{|c|}{Time} & \multicolumn{1}{|c|}{$\mathcal{F}_{\langle p \rangle}$} & \multicolumn{1}{|c|}{Time}\\\hline
1  & Acrobot  & 4 &  & \xmark &  & ~\xmark & & ~\xmark &  & ~\xmark \\
2  & Antenna  & 6 &  & ~\xmark &  & ~\xmark & & ~\xmark & & ~\xmark\\
3  & Ballmaglev  & 3 &  & ~\xmark &  & ~\xmark & & ~\xmark & & ~\xmark\\
4  & Bioreact & 3 & 16,8 & 11.2\,s & 16,10 &14.6\,s & 24,14 & 0.4\,s & 16,10 & 0.5\,s\\
5  & Chen\_ex1& 3 & 16,8 & 7.2\,s  & 16,10 & 9.2\,s & & ~\xmark & 16,10 & 1.4\,s\\
6  & Chen\_ex2 & 8& & ~\xmark & 32,23 & 4683.8\,s &  & ~\xmark & & ~\xmark\\
7  & Chen\_ex3 & 7 &   & ~\xmark & & ~\xmark & & ~\xmark & & ~\xmark\\
8  & Chen\_ex4 & 9 & & ~\xmark & & ~\xmark & & ~\xmark & & ~\xmark\\
9  & Cruise    & 1 & 16,8 & 6.6\,s & 16,10 & 6.5\,s & 24,14 & 0.3\,s& 16,10 & 0.3\,s\\
10 & CruiseHSCC & 1 & 16,8 & 6.6\,s & 16,10 & 6.6\,s &  24,14 & 0.3\,s & 16,10 & 0.3\,s \\
11 & Cstr & 1 & & ~\xmark & & ~\xmark & & ~\xmark & & ~\xmark\\
12 & Cstrtmp  & 1 &  & ~\xmark & & ~\xmark & 24,14 & 0.5\,s & & ~\xmark\\
13 & DCmotor   & 2 & 16,8  & 6.9\,s & 16,10 & 7.8\,s & 24,14 & 0.4\,s & 16,10 & 0.5\,s \\
14 & Flexbeam   & 6 & & ~\xmark & & ~\xmark & 24,14 & 72.1\,s & & ~\xmark\\
15 & Guidance chen  & 5 & & ~\xmark & & ~\xmark & & ~\xmark & & ~\xmark\\
16 & Helicopter   & 3 & 16,8 & 13.8\,s & 32,23 & 148\,s & 24,14 & 0.8\,s& 16,10 & 1.1\,s\\
17 & Invpendulum\_cartpos & 4 & & ~\xmark & & ~\xmark & & ~\xmark & & ~\xmark\\
18 & Invpendulum\_pendang & 2 & 24,12 & 9.8\,s & 16,10 & 28.2\,s & 24,14 & 0.4\,s & 16,10 & 0.5\,s\\
19 & Magpointer   & 2 & 16,8 & 10.1\,s & 16,10 & 20.4\,s & 24,14  & 1\,s& 16,10 & 1.3\,s\\
20 & Magsuspension  & 2 & 12,12  & 22.2\,s  & 16,10 & 34.8\,s & 24,14 & 0.5\,s & 16,10 & 0.6\,s\\
21 & Pendulum   & 2 & 16,8 & 8.5\,s & 32,23 & 24\,s & 24,14 & 0.4\,s & 16,10 & 0.5\,s\\
22 & Regulator   & 5 & & ~\xmark & & ~\xmark & & ~\xmark & & ~\xmark\\
23 & Springmassdamper & 2 & & ~\xmark & & ~\xmark & & ~\xmark & & ~\xmark\\
24 & streamdrum   & 3 & 16,8  & 332.5\,s & 16,10 & 590.5\,s & 24,14 & 0.7\,s & 16,10 &1.3\,s\\
25 & Suspension  & 4 & 16,8  & 17.2\,s & 16,10 &1373\,s & 56,30 & 15.3\,s & 16,10 & 4.4\,s\\
26 & Tapedriver   & 3 & 16,8  & 8.2\,s & & ~\xmark & 64,22 & 2.1\,s & 16,10 & 1.1\,s\\
\hline
%
\end{tabular}
\vspace{0.05in}
\caption{Experimental results\label{tab:results}}
\end{table*}

>>>>

%% \addtodo{CD: the novelty compared to CAV'17 needs to be clarified}
This paper is an extension of~\cite{DBLP:conf/cav/AbateBCCDKKP17}
and makes the following contributions (the fourth and the fifth are
specific extensions over~\cite{DBLP:conf/cav/AbateBCCDKKP17}): 
%
\begin{enumerate}

  % Additions from CAV:
  % continuous time
  % floating point 
\item We automatically generate \emph{correct-by-construction} digital
  controllers using an inductive synthesis approach (CEGIS).  
  This application is non-trivial and addresses challenges specific to digital control
  systems, such as the presence of A/D quantizers and FWL operations.  %% In particular, we
  %% have found that a two-stage verification engine that continuously refines
  %% the precision of the fixed-point representation of the plant yields a
  %% speed-up of two orders of magnitude over a conventional one-stage
  %% verification engine.
%
  %\item  We compute
  The automatically computed state-feedback controllers 
  guarantee a given safety specification alongside a stability property. 
  \note{[this statement is way too broad] Existing methods for controller synthesis rely on transfer 
  function representations, which are inadequate to prove safety requirements. } 
%
\item We provide two novel algorithms: the first, a na\"ive approach, relies on 
  an unfolding of the dynamics up to a completeness threshold, while the
  second one is abstraction-based and leverages abstraction/refinement and
  acceleration to improve scalability while retaining soundness.  %% A third
  %% algorithm evaluates systems using transfer functions where no state-space
  %% description is given.
  Both approaches provide sound synthesis of 
  feedback controllers and consider the various sources of imprecision in
  the implementation of the control algorithm and in the modeling of the
  plant. 
%
\item We develop a model for different sources of quantization error and account for 
  their effect on safety.  We give bounds that ensure the
  safety of our controllers in a hybrid (i.e., continuous/digital) domain.

\item While~\cite{DBLP:conf/cav/AbateBCCDKKP17} is restricted to models in discrete time, in the current paper we present an approach for
  soundly evaluating safety in combined continuous-discrete time and space.

\item A further limitation of the work in~\cite{DBLP:conf/cav/AbateBCCDKKP17} 
  is its restriction to fixed-point arithmetic, meaning that fixed-point
  numbers are employed for the representation of both the plant and the
  controller, as well as for the operations performed by each of them. 
  Conversely, in the current paper we also make use of floating-point
  arithmetic, and compare the two approaches.  Additionally, we enhance the
  implementation with an SMT back-end, allowing us to also use reals in order
  to represent the states of the plant.

@@

\textbf{CEGIS --} Program synthesis is the problem of computing
correct-by-design programs from high-level specifications. Algorithms
for this problem have made substantial progress in recent years, in
particular CEGIS, which is a recent approach to inductive synthesis
borrowing ideas from Counter-Example Guided Abstraction Refinement
(CEGAR)~\cite{DBLP:conf/cav/ClarkeGJLV00} and
superoptimisation~\cite{DBLP:conf/iclp/BrainCVF06,
  DBLP:conf/asplosss/Massalin87}.  
\note{might be confusing that we relate the CEGAR architecture also to Abstract Acceleration?}   
This approach uses an iterative
process \note{cf. Fig. XX}, where each iteration performs inductive generalisation based
on counterexamples provided by a verification oracle.  In the absence
of complete specifications, oracles are queried to distinguish between
non-equivalent
programs~\cite{jha-icse10,DBLP:conf/fmcad/AlurBJMRSSSTU13}.

@@

Program synthesizers are an ideal fit for the synthesis of digital controllers, since
the semantics of programs precisely capture the effects of finite-precision arithmetic.  
In~\cite{DBLP:conf/cdc/RavanbakhshS15}, the authors use CEGIS 
for the synthesis of switching controllers for stabilizing continuous-time
plants with polynomial dynamics.  The work extends to affine systems, but is
limited by the capacity of the state-of-the-art SMT solvers for solving
linear arithmetic.  Since this approach uses switching models instead of
linear dynamics for the digital controller, it avoids problems related to
finite precision arithmetic, but potentially suffers from state-space
explosion.  Moreover, in \cite{DBLP:conf/emsoft/RavanbakhshS16} the same
authors use a CEGIS-based approach for synthesizing continuous-time
switching controllers that guarantee \emph{reach-while-stay} properties of
closed-loop systems, i.e., properties that specify a set of goal states and
safe states (constrained reachability).  This solution is based on
synthesizing control Lyapunov functions for switched systems that yield
switching controllers with a guaranteed minimum dwell time in each mode. 
However, both approaches are unsuitable for the kind of controllers that we seek to
synthesize.

The work in~\cite{hscc-paper} synthesizes stabilizing controllers for
continuous plants given as transfer functions by exploiting bit-accurate
verification of software-implemented digital controllers~\cite{Bessa16}. 
While this work also uses CEGIS, the approach is restricted to digital
controllers for stable closed-loop systems given as transfer function
models: this results in a static check on their coefficients.  By contrast,
in the current paper we consider a state-space representation of the
physical system, which requires ensuring the specification over actual
traces of the model, alongside the numerical soundness required by the
effects of discretisation and finite-precision errors.  
\note{Next paragraph is naive for a control audience - amend} 
A~state-space model
has known advantages over the transfer function
representation~\cite{Franklin15}: it naturally generalizes to multivariate
systems (i.e., with multiple inputs and outputs); and it allows synthesis of
control systems with guarantees on the internal dynamics, e.g., to
synthesize controllers that make the closed-loop system \emph{safe}.  Our
work focuses on the \emph{safety} of internal states, which is usually
overlooked in the literature.  Moreover, our work integrates an
abstraction/refinement (CEGAR) step inside the main CEGIS loop.

The tool Pessoa~\cite{mazo2010pessoa} synthesizes correct-by-design embedded
control software in a Matlab toolbox.  It is based on the abstraction of a
physical system to an equivalent finite-state machine and on the computation
of reachability properties thereon.  Based on this safety specification,
\mbox{Pessoa} can synthesize embedded controller software for a range of
properties.  The embedded controller software can be more complicated than
the state-feedback control we synthesize, and the properties available cover
more detail.  However, relying on state-space discretization \mbox{Pessoa}
is likely to incur in scalability limitations.  \note{Mention SCOTS?}
Along this research line,
\cite{Anta2010,liu16} studies the synthesis of digital controllers for
continuous dynamics, and \cite{zamani2014} extends the approach to the Network Control Systems setup.

\textbf{Discretization Effects in Control Design --}
Classical work in control synthesis has often disregarded digitalization effects, 
\note{[this paragraph needs more work:]
whereas recent results in digital control have focused on separate aspects of discretization, 
including delayed response~\cite{Duggirala2015} and finite word length (FWL) semantics, 
with the goal either to verify~\cite{daes20161} or to optimize~\cite{oudjida2014design} given implementations. }

@@

\begin{table*}
\centering
\begin{tabular}{| r | l | c | c | r | c | r | c | r | c | r |}
%
\hline
\# & \multicolumn{1}{|c|}{Benchmark} & \multicolumn{1}{|c|}{Order} & \multicolumn{4}{|c|}{Multi-staged}                 & \multicolumn{4}{|c|}{Abstraction} \\
   &                                  & & \multicolumn{1}{|c|}{$\mathcal{F}_{\langle I_p,F_p \rangle}$} & \multicolumn{1}{|c|}{Time} & \multicolumn{1}{|c|}{$\mathcal{F}_{\langle p \rangle}$} & \multicolumn{1}{|c|}{Time} & \multicolumn{1}{|c|}{$\mathcal{F}_{\langle I_p,F_p \rangle}$} & \multicolumn{1}{|c|}{Time} & \multicolumn{1}{|c|}{$\mathcal{F}_{\langle p \rangle}$} & \multicolumn{1}{|c|}{Time}\\\hline
1  & Acrobot  & 4 &   &  &  &  & & & &\\
2  & Antenna  & 6 &   &  &  &  & & & &\\
3  & Ballmaglev  & 3 &  &   &  &  & & & &\\
4  & Bioreact & 3 & 16,8 & 11200ms & 16,10 &14600ms & 24,14 & 425ms& 16,10 & 502ms\\
5  & Chen\_ex1& 3 & 16,8 & 7278ms  & 16,10 & 9196ms & & & 16,10 & 1398ms\\
6  & Chen\_ex2 & 8& & & 32,23 & 4683872ms &  & & &\\
7  & Chen\_ex3 & 7 &   & & & & & & &\\
8  & Chen\_ex4 & 9 &  & & & & & & &\\
9  & Cruise    & 1 & 16,8 & 6563ms & 16,10 & 6543ms & 24,14 & 295ms& 16,10 & 306ms\\
10 & CruiseHSCC & 1 & 16,8 & 6587ms & 16,10 & 6567ms &  24,14 &293ms& 16,10 & 313ms\\
11 & Cstr & 1 & &  & & & & & &\\
12 & Cstrtmp  & 1 &   &  & & & 24,14 &478ms& &\\
13 & DCmotor   & 2 & 16,8  & 6879ms & 16,10 & 7786ms & 24,14 & 435ms& 16,10 & 520ms\\
14 & Flexbeam   & 6 &   &  & & & 24,14 &72079ms& &\\
15 & Guidance chen  & 5 &   &  & & & & & &\\
16 & Helicopter   & 3 & 16,8 & 13728ms & 32,23 & 148012ms & 24,14 & 727ms& 16,10 &1075ms\\
17 & Invpendulum\_cartpos & 4 &   &  & & & & & &\\
18 & Invpendulum\_pendang & 2 & 24,12 & 9741ms & 16,10 & 28135ms & 24,14 & 432ms& 16,10 &466ms\\
19 & Magpointer   & 2 & 16,8 & 10093ms & 16,10 & 20407ms & 24,14  & 916ms& 16,10 &1244ms\\
20 & Magsuspension  & 2 & 12,12  &22146ms  & 16,10 & 34808ms & 24,14 & 443ms & 16,10 &569ms\\
21 & Pendulum   & 2 & 16,8 & 8503ms & 32,23 & 23987ms & 24,14 & 395ms& 16,10 &473ms\\
22 & Regulator   & 5 &   &  & & & & & &\\
23 & Springmassdamper & 2 &   &  & & & & & &\\
24 & streamdrum   & 3 & 16,8  & 332422ms & 16,10 & 590501ms & 24,14 & 701ms& 16,10 &1225ms\\
25 & Suspension  & 4 & 16,8  & 17206ms & 16,10 &1372945ms & 56,30 & 15268ms& 16,10 &4374ms\\
26 & Tapedriver   & 3 & 16,8  & 8188ms & & & 64,22 & 2026ms& 16,10 &1090ms\\
\hline
%
\end{tabular}
\vspace{0.05in}
\caption{Experimental results\label{tab:results}}
\end{table*}
